Neural Connectome of the Ctenophore Statocyst

### 2025-06-02 manuscript/Manuscript.qmd
html

![**Figure 3. Synaptic connectivity of the balancer and and bridge cells with the ANN.**\
(A) Number of synaptic inputs from the ANNs (ANN_Q1-4: blue, ANN_Q1Q2: orange, ANN_Q3Q4: pink) to each cell type. Abbreviations: bal, balancer; brg, bridge; bsl, bristle; cg, ciliated groove; dv, dense vesicle cell; imc, intra-multiciliated cell; la, lamellate body; li, lithocyte; pl, plumose cell; ef, epithelial floor cell.\
<!-- check order of panels, change here or change figure --> (B) Connectivity map of three ANN neurons (Q1-4 in blue, Q1Q2 in pink, and Q3Q4 in orange) and balancer cells (light blue). The thickness of the arrows and the numbers correspond to the number of synapses. Synaptic structures from the same neuron targeting the same balancer cell are grouped into hexagons, with the number of cells added to the label.\
(C) Morphological rendering of bridge cells spanning the Q1Q2 and Q3Q4 quadrants. The Q1Q2-side bridge cells (8 cells) are shown in blue, while the Q3Q4-side bridge cells (6 cells) are shown in xxx color. The morphology of individual bridge cells extending across opposite quadrant regions is depicted. The spheres represent the positions of individual nuclei. The left panel shows an aboral view of the aboral organ, the middle panel presents a sagittal plane view, and the right panel provides a tentacular plane view.\
(D) Mitochondrial localization within bridge cells and associated presynaptic triad structures. Red indicates mitochondria associated with presynaptic triad structures, yellow marks mitochondria containing synaptic vesicles but lacking a clearly defined presynaptic triad, blue represents mitochondria with unclear synaptic vesicles, and black denotes mitochondria where no synaptic vesicles were identified. The left panel shows a dorsal view of the aboral organ, the middle panel presents a sagittal plane view, and the right panel provides a tentacular plane view.\
(E) Synaptic connections from ANN neurons to balancer ciliated cells. The positions of synapses from ANNs to balancers (magenta) are indicated. The left panel presents an aboral view, while the right panel shows a lateral view and a tentacular plane view. Balancer ciliated cells are depicted in light gray.\
(F) Synaptic connections between ANNs and bridge cells. The positions of synapses from ANNs to bridge cells (magenta) and from bridge cells to ANNs (light blue) are indicated. The left panel shows an aboral view, while the right panel presents a tentacular plane lateral view. Bridge cells are shown in light gray.\
(G) Connectivity matrix of the gravity-sensing neural circuit. Columns represent presynaptic cell groups, while rows represent postsynaptic cell groups. The numbers and varying shades of blue correspond to the number of synapses.\
(H) Complete connectivity map of the gravity-sensing neural circuit. Cells belonging to the same group are enclosed in hexagons, and the number of cells is added to their labels. The thickness of the arrows and the numerical values indicate the number of synapses. ANN Q1-4 are shown in blue, ANN Q1Q2 in orange, ANNs in Q3Q4 in pink, balancer cell groups in gray, and bridge cell groups in yellow.](figures/Figure3.png)

@@

![**Video 1.** A...]

![**Video 2.** A...]

![**Video 3.** This video shows high-speed imaging of balancer ciliary beating along with kymographs from two individual balancer cilia. The recording was acquired at 100 frames per second and played back at 30 frames per second (0.3×). The white line overlaid on the kymographs indicates the current time point shown in the video. This example captures a ciliary arrest event in the sagittal plane.]

![**Video 4.**This video shows high-speed imaging of balancer ciliary beating along with kymographs from two individual balancer cilia. The recording was acquired at 100 frames per second and played back at 30 frames per second (0.3×). The white line overlaid on the kymographs indicates the current time point shown in the video. This example captures a ciliary re-beat event in the sagittal plane.]

![**Video 5.** This video shows high-speed imaging of balancer ciliary beating along with kymographs from two individual balancer cilia. The recording was acquired at 100 frames per second and played back at 30 frames per second (0.3×). The white line overlaid on the kymographs indicates the current time point shown in the video. This example captures a ciliary arrest event in the tentacular plane.]

![**Video 6.**This video shows high-speed imaging of balancer ciliary beating along with kymographs from two individual balancer cilia. The recording was acquired at 100 frames per second and played back at 30 frames per second (0.3×). The white line overlaid on the kymographs indicates the current time point shown in the video. This example captures a ciliary re-beat event in the tentacular plane.]

>>>>

![**Figure 3. Synaptic connectivity of the balancer and and bridge cells with the ANN.**\
(A) Number of synaptic inputs from the ANNs (ANN_Q1-4: blue, ANN_Q1Q2: orange, ANN_Q3Q4: pink) to each cell type. Abbreviations: bal, balancer; brg, bridge; bsl, bristle; cg, ciliated groove; dv, dense vesicle cell; imc, intra-multiciliated cell; la, lamellate body; li, lithocyte; pl, plumose cell; ef, epithelial floor cell.\
(B) Morphological rendering of balancer cells, with each cell color-coded by quadrant: Q1 (yellow), Q2 (blue), Q3 (orange), and Q4 (magenta). Each sphere represents the position of an individual nucleus. The fine projections extending from the cells represent traced trajectories that follow the cell body and continue into a single cilium from the point where it emerges. Views are shown from three orthogonal perspectives: aboral view (left), sagittal plane (middle), and tentacular plane (right). The spatial arrangement highlights the quadrant-specific organization of balancer clusters within the aboral organ. Scale bar: 25 µm.\
(C) Morphological rendering of bridge cells spanning the Q1Q2 and Q3Q4 quadrants. The Q1Q2-side bridge cells (8 cells) are shown in blue, while the Q3Q4-side bridge cells (6 cells) are shown in xxx color. The morphology of individual bridge cells extending across opposite quadrant regions is depicted. The spheres represent the positions of individual nuclei. The left panel shows an aboral view of the aboral organ, the middle panel presents a sagittal plane view, and the right panel provides a tentacular plane view.\
(D) Mitochondrial localization within bridge cells and associated presynaptic triad structures. Red indicates mitochondria associated with presynaptic triad structures, yellow marks mitochondria containing synaptic vesicles but lacking a clearly defined presynaptic triad, blue represents mitochondria with unclear synaptic vesicles, and black denotes mitochondria where no synaptic vesicles were identified. The left panel shows a dorsal view of the aboral organ, the middle panel presents a sagittal plane view, and the right panel provides a tentacular plane view.\
(E) Synaptic connections from ANN neurons to balancer ciliated cells. The positions of synapses from ANNs to balancers (magenta) are indicated. The left panel presents an aboral view, while the right panel shows a lateral view and a tentacular plane view. Balancer ciliated cells are depicted in light gray.\
(F) Synaptic connections between ANNs and bridge cells. The positions of synapses from ANNs to bridge cells (magenta) and from bridge cells to ANNs (light blue) are indicated. The left panel shows an aboral view, while the right panel presents a tentacular plane lateral view. Bridge cells are shown in light gray.\
(G) Connectivity matrix of the gravity-sensing neural circuit. Columns represent presynaptic cell groups, while rows represent postsynaptic cell groups. The numbers and varying shades of blue correspond to the number of synapses.\
(H) Complete connectivity map of the gravity-sensing neural circuit. Cells belonging to the same group are enclosed in hexagons, and the number of cells is added to their labels. The thickness of the arrows and the numerical values indicate the number of synapses. ANN Q1-4 are shown in blue, ANN Q1Q2 in orange, ANNs in Q3Q4 in pink, balancer cell groups in gray, and bridge cell groups in yellow.](figures/Figure3.png)

@@

**Video 1.** This video shows a rotating 3D rendering of the aboral organ, featuring three distinct cell types—balancer cells, bridge cells, and components of the apical nerve net (ANN). Cells of each type are subdivided and color-coded according to their spatial regions. A lithocyte is shown in gray as a reference point, and the outline of the aboral organ is indicated in white.

**Video 2.** This video shows high-speed imaging of balancer ciliary beating along with kymographs from two individual balancer cilia. The recording was acquired at 100 frames per second and played back at 30 frames per second (0.3×). The white line overlaid on the kymographs indicates the current time point shown in the video. This example captures a ciliary arrest event in the sagittal plane.

**Video 3.** This video shows high-speed imaging of balancer ciliary beating along with kymographs from two individual balancer cilia. The recording was acquired at 100 frames per second and played back at 30 frames per second (0.3×). The white line overlaid on the kymographs indicates the current time point shown in the video. This example captures a ciliary re-beat event in the sagittal plane.

**Video 4.** This video shows high-speed imaging of balancer ciliary beating along with kymographs from two individual balancer cilia. The recording was acquired at 100 frames per second and played back at 30 frames per second (0.3×). The white line overlaid on the kymographs indicates the current time point shown in the video. This example captures a ciliary arrest event in the tentacular plane.

**Video 5.** This video shows high-speed imaging of balancer ciliary beating along with kymographs from two individual balancer cilia. The recording was acquired at 100 frames per second and played back at 30 frames per second (0.3×). The white line overlaid on the kymographs indicates the current time point shown in the video. This example captures a ciliary re-beat event in the tentacular plane.
